---
title: "lab8"
author: "Gabrielle"
date: '2022-11-14'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = FALSE)
library(tidyverse)
library(here)
library(ggpubr)
library(janitor)
library(broom)
library(knitr)
```

# lab 8: dummy variables, tables, and model selection using BIC

### load in the data
```{r}
penguins <- read_csv(here("data", "penguins.csv"))
```


### examples of pairs() and cor()
```{r, results = TRUE}
penguins_mod <- penguins %>% 
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g) %>% 
  drop_na()
# selecting the continuous variables, and not the discrete ones, because correlation wouldn't look right for discrete variables

pairs(penguins_mod)
cor(penguins_mod)
```

## Linear models
```{r}
# new data set
penguins_Adelie_Torgersen <- penguins %>% 
  filter(species == "Adelie") %>% 
  filter(island == "Torgersen") %>% 
  drop_na()
```

## models
```{r}
model_0 <- lm(body_mass_g ~ 1, data = penguins_Adelie_Torgersen)
# a regression without anything in "x" it's the mean of the variable y

model_1 <- lm(body_mass_g ~ flipper_length_mm, data = penguins_Adelie_Torgersen)
# body mass is a function of flipper length

model_2 <- lm(body_mass_g ~ sex, data = penguins_Adelie_Torgersen)
# body mass is a function of sex

model_3 <- lm(body_mass_g ~ flipper_length_mm + sex, data = penguins_Adelie_Torgersen)
# body mass is a function of flipper length coupled with the sex



## model outputs saved to data

model_0_out <- summary(model_0)

model_1_out <- summary(model_1)

model_2_out <- summary(model_2)

model_3_out <- summary(model_3)

### now we can start using the model results and make data tables

```


Model 0 This null model captures the idea that there is no explanatory power of flipper length or sex. It is useful for comparison to other models with more complexity.

Model 1 This model is a simple linear regression.

Model 2 This model with only a dummy variable for sex.

Model 3 This model with a dummy variable for sex and a continuous variable of flipper length.


**Model 0**
```{r, results = TRUE}
out_0_tidy <- tidy(model_0)
out_0_tidy
# 'straw man stupid test'... is the penguin mass different than zero... well yes, of course it is

out_0_glance <- glance(model_0)
out_0_glance
# r squared is zero because there is only one variable
# sigma, variance of the residuals
# we're not testing a regression model, so many of the results are NA
# there is a model being run, so you can see the logLik and AIC and BIC

kable(out_0_tidy, format = "markdown", digits = 3, caption = "Tests of linear model (model 0) coefficients")

kable(out_0_glance, format = "markdown", digits = 3, caption = "Tests of linear model (model 0)")

```


**Model 1**
```{r, results = TRUE}
out_1_tidy <- tidy(model_1)
out_1_tidy

out_1_glance <- glance(model_1)
out_1_glance

kable(out_1_tidy, format = "markdown", digits = 3, caption = "Tests of linear model (model 1) coefficients")

kable(out_1_glance, format = "markdown", digits = 3, caption = "Tests of linear model (model 1)")

```



**Model 2**
```{r, results = TRUE}
out_2_tidy <- tidy(model_2)
out_2_tidy

out_2_glance <- glance(model_2)
out_2_glance

kable(out_2_tidy, format = "markdown", digits = 3, caption = "Tests of linear model (model 2) coefficients")

kable(out_2_glance, format = "markdown", digits = 3, caption = "Tests of linear model (model 2)")

```



**Model 3**
```{r, results = TRUE}
out_3_tidy <- tidy(model_3)
out_3_tidy

out_3_glance <- glance(model_3)
out_3_glance

kable(out_3_tidy, format = "markdown", digits = 3, caption = "Tests of linear model (model 3) coefficients")

kable(out_3_glance, format = "markdown", digits = 3, caption = "Tests of linear model (model 3)")


```






